---
title: "Homework 8"
author: "Carmen Canedo"
date: "27 October 2020"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r, message=FALSE}
source("setup.R")
```


# Exercise 8
Create regular expressions to find all words that:

  + End with "ed" but not "eed"
  + End with "ing" or "ise"
  + Words that do not follow the rule "i before e except after c"
  
```{r}
# Words that end with 'ed', but not 'eed'
words %>% 
  str_view("[^e]ed$", match = TRUE)
```

```{r}
# End with -ing or -ise
words %>% 
  str_view("(ing|ise)$", match = TRUE)
```

```{r}
# Doesn't follow 'i before e'
words %>% 
  str_view(".*cei.*", match = TRUE)
```

# Exercise 9
Create regular expressions to find all words that:

  + start or end with the letter x
  + start with at least two vowels and end with at least two consonants
  + Contains three or more vowels in a row
  
```{r}
# Starts or end with x
words %>% 
  str_view("x^|x$", match = TRUE)
```
  
```{r}
# Starts with at least two vowels
# Ends with at least two consonants
words %>% 
  str_view("^[aeiouy]{2,}.*[^aeiouy]${2,}", match = TRUE)
```
  
```{r}
# Contains three or more vowels in a row
words %>% 
  str_view("[aeiouy]{3}", match = TRUE)
```

# Exercise 10
Construct regular expressions to find words that:

  + Start and end with the same character
  + Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice)
  + Contain one letter other than e that is repeated in at least three places (e.g. appropriate contains three "p's")
  
```{r}
words %>% 
  str_view("", match = TRUE)
```

```{r}

```

