---
title: "Lab 1"
author: "Carmen Canedo"
date: "6 October 2020"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: inline
---

```{r Setup File, include=FALSE, message=FALSE}
# Loading libraries
source("setup.R")
```

# Exercise 1
Read in `avocadoes.csv` and save into data fram.
```{r Reading Avocados .csv}
# Reading in avocados.csv
avocados <- read_csv("avocados.csv")
```

# Exercise 2
Change column names to small, large, and extra large.
```{r Rename Avo Cols}
# Renaming columns to appropriate size
avocados <- avocados %>% 
  rename(small = `4046`,
         large = `4225`,
         extra_large = `4770`)
```

# Excercise 3
The data is not tidy, I will need to use `pivot_longer()` to tidy it. The size variable is split into small, large, and extra large.
```{r}
avocados <- avocados %>% 
  pivot_longer(cols = c(small, large, extra_large), names_to = "size", values_to = "sales_volume")
```


# Exercise 4
Create a function to read in the avocado data, renames columns, writes the transformed, and returns transformed data
```{r Function to read in avo .csv}
load_clean_data <- function(file, subdirectory_name) {
  
  # Read in the avocado data
  tbl <- read_csv(file)
  
  # Renames avocado size columns
  tbl <- tbl %>% 
    rename(small = `4046`,
         large = `4225`,
         extra_large = `4770`)
  
  # Tidies data
  tbl <- tbl %>% 
    pivot_longer(cols = c(small, large, extra_large), names_to = "size", values_to = "sales_volume")
  
  # Check if directory exists, creates it if not
  n_subdirectories <- length(subdirectory_name)
  
  for (i in 1:n_subdirectories) {
    if (!dir.exists(subdirectory_name[i])) {
      dir.create(subdirectory_name[i])
    }
  }
  
  # Write .csv of file in specified location
  data_location <- paste(subdirectory_name, "/", file, sep = "")
  
  #  Check file doesn't exist at location
  if(!file.exists(data_location)) {
      write_csv(tbl, path = data_location)
  }
  
  tbl_final <- read_csv(data_location)
  
  return(tbl_final)
}
```

```{r Confirm Function Works}
avocados_test <- load_clean_data("avocados.csv", "avo_data")
```

# Exercise 5
Create table with top 5 geographies with highest total yearly sales of large avocados in 2019
```{r Kable Table}
# Creating table
avocados %>% 
  # Showing only 2019
  filter(year == 2019) %>% 
  # Selecting only necessary columns
  select(large, geography) %>%
  # From highest to lowest
  arrange(desc(large)) %>% 
  # Rename columns
  rename(`Total Yearly Sales` = large,
         `Location` = geography) %>% 
  # Show top 5
  head(n = 5) %>% 
  # Make table
  kable()
```

# Exercise 6
Create a plot showing daily sales volume in California for different avocado types and sizes.
```{r Cali Scatterplot 1}
# Subsetting data with only necessary info
cali_avo <- avocados %>% 
  filter(geography == "California") %>% 
  select(-year, -average_price)

# Marking base scatterplots for each size
small <- cali_avo %>% 
  ggplot(aes(x = date, y = small)) +
  geom_point() +
  geom_jitter() +
  facet_grid(~ type, scales = "free_y")

large <- cali_avo %>% 
  ggplot(aes(x = date, y = large)) +
  geom_point() +
  geom_jitter() +
  facet_grid(~ type, scales = "free_y")

extra_large <- cali_avo %>% 
  ggplot(aes(x = date, y = extra_large)) +
  geom_point() +
  geom_jitter() +
  facet_grid(~ type, scales = "free_y")
```

```{r Cali Scatterplot 2}
# Adding details to each graph
small <- small +
  labs(title = "Daily Sales Volume in California",
       subtitle = "Size: Small",
       x = "Date",
       y = "Number of avocados sold") +
  theme_light()

large <- large +
    labs(title = "Daily Sales Volume in California",
       subtitle = "Size: Large",
       x = "Date",
       y = "Number of avocados sold") +
  theme_light()

extra_large <- extra_large +
    labs(title = "Daily Sales Volume in California",
       subtitle = "Size: Extra Large",
       x = "Date",
       y = "Number of avocados sold") +
  theme_light()
```


```{r Printing Cali Scatterplots}
# Printing graphs
small
large
extra_large
```

# Exercise 7
Load `heart.txt`
```{r Read in heart data}
# Loading data
heart <- read_delim("heart.txt",
                    delim = "\t" ,
                     col_types = cols(
                       anaemia = readr::col_factor(levels = c("Yes", "No")),
                       DEATH_EVENT = readr::col_factor(levels = c("Yes", "No")),
                       sex = readr::col_factor(levels = c("Male", "Female")),
                       smoking = readr::col_factor(levels = c("Yes", "No")),
                       high_blood_pressure = readr::col_factor(levels = c("Yes", "No")),
                       diabetes = readr::col_factor(levels = c("Yes", "No"))
                    )
                    )
```

# Exercise 8
```{r Heart boxplot 1}
# Creating labels to rename facet titles
smoker_labs <- c("Non-Smoker", "Smoker")
names(smoker_labs) <- c("No", "Yes")

# Making base boxplot
heart_box <- heart %>% 
  ggplot(aes(x = DEATH_EVENT, y = ejection_fraction)) +
  geom_boxplot(aes(fill = anaemia)) +
  facet_wrap(~ smoking, labeller = labeller(smoking = smoker_labs))
```

```{r Heart boxplot 2}
# Adding text
heart_box <- heart_box +
  labs(title = "Event of Death from Ejection Fraction",
       x = "Event of Death",
       y = "Ejection Fraction")
```


```{r Print boxplots}
# Printing final graph
heart_box
```

```{r}
### Ran out of time
```

