---
title: "Homework 9"
author: "Carmen Canedo"
date: "28 November 2020"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r, message=FALSE}
# Loading librarie
source("setup.R")
library(lubridate)
```


# Exercise 1
## Question 1: Explore the distribution of `rincome` (reported income). What makes the default bar chart hard to understand? How could you improve the plot?
```{r}

```


## Question 2: What is the most common religion? Most common `partyid`?
```{r}

```

## Question 3: Which `relig` does `denom` apply to? How can you find out with a table? How can you find out with a vizualization?
```{r}

```

## Question 4: How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?
```{r}

```

## Question 5: How could you collapse `rincome` into a small set of categories?
```{r}

```


# Exercise 2
## Question 1: Parse the following date-times
"05/26/2004 UTC 11:11:11.444"
"26 2004 05 UTC 11/11/11.444"
```{r}

```

## Question 2: Use the appropriate `lubridate` function to parse each of the following dates.
```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)") d5 <- "12/30/14" # Dec 30, 2014
```

## Question 3: For each month in `wmata_ridership`, calculate the proportion of rides made on a given day of the month. Then make box plots of the proportions of ridership vs. day of the week, *excluding any days from 2004*.
```{r}

```

## Question 4: How long of a time-span is covered in the WMATA ridership dataset?
```{r}

```

