---
title: "Homework 7"
author: "Carmen Canedo"
date: "27 November 2020"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: inline
---

# Loading Libraries
```{r Load Data, message=FALSE, echo=FALSE}
source("setup.R")

library(nycflights13)

nycflights13_data <- data(package = "nycflights13")
```

# Keys
*Primary keys* uniquely identify an observation within its own table. *Foreign keys* uniquely identify an observation in another table.

To be sure the identification is unique, count the primary keys and see if any are greater than 1.

```{r Confirm unique key 1}
planes_test <- planes %>% 
  count(tailnum) %>% 
  filter(n > 1)
```

```{r Confirm unique key 2}
flights_test <- flights %>% 
  count(year, month, day, flight) %>% 
  filter(n > 1)
```
There is not a primary key in the flights tibble because the values repeat. We need to add a *surrogate key* to flights.

```{r Add surrogate key}
flightskey <- flights %>% 
  rowid_to_column("ID") %>% 
  select(ID, everything())
```

## Exercise 1 Identify the primary keys in three datasets. 
Be sure to show that you have the primary key by showing there are no duplicate entries.

### a. Lahman Batting Primary Key
```{r Lahman::Batting PK}
# Loading Library
library(Lahman)

#View(Batting)

# Finding primary key
Batting %>% 
  count(playerID, stint, yearID) %>% 
  filter(n > 1)
```
The primary key for `Lahman::Batting` is `playerID`, `stint`, and `yearID`.

### b. Babynames Primary key
```{r babynames::babynames PK}
library(babynames)

#View(babynames)

# For some reason, when compiling, the code lags here, so I'm commenting it out. This is how I got my answer for this question.
# babynames::babynames %>% 
#   count(name, year, sex) %>% 
#   filter(n > 1)
```
The primary key for `babynames` is `name`, `year`, and `sex`. 

### c. Atmos Primary Key
```{r nasaweather::atmos PK}
nasaweather::atmos %>% 
  count(lat, long, year, month) %>% 
  filter(n > 1)
```
the primary key for `atmos` is `lat`, `long`, `year`, `month`.

## Exercise 2 What is the relationship between the Batting, Master, and Salaries tables in the Lahman package? What are the keys for ech dataset and how do they relate to each other?

```{r Explore Lahman data}
# View(Batting)
# View(Master)
# View(Salaries)
```

```{r Master PK}
Master %>% 
  count(playerID) %>% 
  filter(n > 1)
```

```{r Salaries PK}
Salaries %>% 
  count(playerID, teamID, yearID) %>% 
  filter(n > 1)
```


As we know from above, the primary key for `Batting` is `playerID`, `stint`, and `yearID`. The primary key for `Master` is `playerID`, and for `Salaries` it is `playerID`, `teamID`, and `yearID`.

`playerID` in `Master` has a 1 to many relationship in Salaries and Batting.

# Mutating Joins
*Inner Joins* connect pairs of observations when their keys are equal. Unmatched rows will not be included in the result. 

*Outer Joins* keep observations that appear in at least one of the tables:

  + `left_join(x, y)` keeps all observations in `x`
  + `right_join(x, y)` keeps all observations in `y`
  + `full_join(x, y)` keeps all observations
  
*Natural join* is the default and uses all variables with the same name in both tables.

Using `by = "variable_name` joins by specified variables only.

You can also use `by = c("variable_a" = "variable_b")` to use a key with a different name in each table

## Exercise 3 Use an appropriate join to add a column containing the airline name to the flights dataset.
Be sure to put the carrier code and name in the first two columns of the result. **Save as `flights2`**
```{r}
flights2 <- flights %>% 
  left_join(airlines, by = "carrier") %>% 
  select(carrier, name, everything())
```


## Exercise 4 Use an appropriate join to add the airport name to the `flights2` dataset.
The codes and names of the airports are in the airports dataset of `nycflights13` package.
```{r}
# Getting rid of unnecessary columns
airport_names <- airports %>% 
  select(faa, name) %>% 
  rename(airport = name)

# Joining to flights2
flights2 %>% 
  left_join(airport_names, by = c("origin" = "faa")) %>% 
  left_join(airport_names, by = c("dest" = "faa"),
            suffix = c("_origin", "_dest")) %>% 
  select(carrier, name, origin, airport_origin, dest, airport_dest, everything())
```


# Filtering Joins
You can use *filter joins* to filter observations from one tibble based on whether or not they match an observation in another tibble.

A *semi join* between `x` and `y`:

  + Keeps all observations in `x` that have a match in `y`
  + Is useful for matchinf filtered summary back to original row
  
An *anti join* between `x` and `y` is the inverse of the semi join:

  + *Drops* all observations in `x` that have a match in `y`
  + Useful for diagnosing join mismatches
  
## Exercise 5 Compute the average delay by destination, then join the airports data frame so you can show the spatial distribution of delays.

  + Use the size or color of the points to display the average delay for each airport.
  + Add the location of the origin and destination (lat and lon) to flights
  + Compute the average delay by destination
  
```{r}

```

  
# Set Operations
`dplyr` provides functions for performing standard set operations:

  + `intersect(x, y)` returns the intersection of datasets `x` and `y`
  + `union(x, y)` returns the union of datasets `x` and `y`
  + `setdiff(x, y)` returns set members in `x` that are not in `y`.
  
## Exercise 6 Use a set operation function to find which airport codes from flights are not in the airports datset
```{r}

```

