---
title: "Homework 8"
author: "Carmen Canedo"
date: "27 October 2020"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r, message=FALSE}
source("setup.R")
```


# Exercise 8
Create regular expressions to find all words that:

  + End with "ed" but not "eed"
  + End with "ing" or "ise"
  + Words that do not follow the rule "i before e except after c"
  
```{r}
# Words that end with 'ed', but not 'eed'
words %>% 
  str_view("[^e]ed$", match = TRUE)
```

```{r}
# End with -ing or -ise
words %>% 
  str_view("(ing|ise)$", match = TRUE)
```

```{r}
# Doesn't follow 'i before e'
words %>% 
  str_view(".*cei.*", match = TRUE)
```

# Exercise 9
Create regular expressions to find all words that:

  + start or end with the letter x
  + start with at least two vowels and end with at least two consonants
  + Contains three or more vowels in a row
  
```{r}
# Starts or end with x
words %>% 
  str_view("x^|x$", match = TRUE)
```
  
```{r}
# Starts with at least two vowels
# Ends with at least two consonants
words %>% 
  str_view("^[aeiouy]{2,}.*[^aeiouy]${2,}", match = TRUE)
```
  
```{r}
# Contains three or more vowels in a row
words %>% 
  str_view("[aeiouy]{3}", match = TRUE)
```

# Exercise 10
Construct regular expressions to find words that:

  + Start and end with the same character
  + Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice)
  + Contain one letter other than e that is repeated in at least three places (e.g. appropriate contains three "p's")
  
```{r}
words %>% 
  str_subset("^(.)((.*\\1$)|\\1?$)")
```

```{r}
words %>% 
  str_subset("([A-Za-z][A-Za-z]).*\\1")
```

```{r}
words %>% 
  str_subset("([a-z]).*\\1.*\\1")
```

# Exercise 11
Use `str_detect()` to find any words that contain at least one of each different vowel.
```{r}
pattern <-
  cross(rerun(5, c("a", "e", "i", "o", "u")),
    .filter = function(...) {
      x <- as.character(unlist(list(...)))
      length(x) != length(unique(x))
    }
  ) %>%
  map_chr(~str_c(unlist(.x), collapse = ".*")) %>%
  str_c(collapse = "|")

pattern
```

```{r}
str_subset(words, pattern)

words[str_detect(words, "a") &
  str_detect(words, "e") &
  str_detect(words, "i") &
  str_detect(words, "o") &
  str_detect(words, "u")]
```


# Exercise 12
Switch the first and last letters in each word of words. Which of those strings are still words?
```{r}
# Switching all words
swap <- str_replace_all(words, "^([A-Za-z])(.*)([A-Za-z])$", "\\3\\2\\1")
```


```{r}
# Finding ones that are still actual words
str_replace_all(words, "^([A-Za-z])(.*)([A-Za-z])$", "\\3\\2\\1") %>% 
  intersect(words)
```

# Exercise 13
Modify the regex so it no longer picks up the word flickered as a color.
```{r}
# Regex to look for colors
colors <- c("red", "Red","orange", "yellow", "green", "blue", "purple")
color_match <- str_c(colors, collapse = "| ")
```

```{r}
# Sentences with more than 1 color
# Select the sentences that contain a color, and then extract the color to figure out which one it is
has_color <- str_subset(sentences, color_match)
matches <- str_extract(has_color, color_match)
# Select all the sentences that have more than 1 match
more <- sentences[str_count(sentences, color_match) > 1]
str_view_all(more, color_match)
```
To make sure the regex doesn't highlight "flickered" our `str_c()` should be modified to the following...
```{r}
# New regex pattern
color_match <- str_c(str_c("\\b", colors, "\\b"), collapse = "|")

# New result
str_view_all(more, color_match)
```


  1. Show all the sentences that contain at least one color with the word highlighted using `str_view_all()`.
```{r}
has_color <- sentences[str_count(sentences, color_match) >= 1]
str_view_all(has_color, color_match)
```

  2. Use `str_extract_all()` to produce an array with the colors that appear multiple times in a sentence
```{r}
has_color <- sentences[str_count(sentences, color_match) > 1]
str_extract_all(has_color, color_match)
```
  
  
# Exercise 14
Find all words that come after a "number" like "one", "two",... "twelve". Pull out both the number and the word.
```{r}
# Regex pattern
numbers <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"

# Extract the words
sentences[str_detect(sentences, numbers)] %>%
  str_extract(numbers)
```

# Exercise 15

  1. View all sentences that contain the word "good" or the word "bad". Get the sentence numbers where those occur.
```{r}
# Pattern
words_to_find <- " good | bad "

# Matching in dataset
sentences[str_detect(sentences, words_to_find)] %>% 
  str_view_all(words_to_find)
```

  2. Use `str_replace_all()` to replace the word "bad" with "horrible and the word "good" with "great". Look at the sentence numbers you found before to verify the words were replaced correctly.
```{r}
good <- str_replace_all(sentences, " good ", " great ")
bad <- str_replace_all(sentences, " bad ", " horrible ")
```
  
  
# Exercise 16
Install `ISLR` package. Then take `Names <- row.names(College)`. Show the first twenty rows that result if you split `Names` at each space into 3 parts. Do the same by splitting the first 20 rows on every word boundary.