---
title: "Lab 2"
author: "Carmen Canedo"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: inline
---

```{r Loading Libraries, message=FALSE}
# Load libraries
source("setup.R")
library(lubridate)
```


# Question 1:
Read in `us-counties.csv` and create a tibble data frame that contains both deaths and total cases per state, arranged by total number of deaths in descending order.
```{r Load US data, message=FALSE}
# Loading data
us_counties <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")
```

```{r Q1 US dataframe, message=FALSE}
# Total cases and deaths by state, arranged by deaths
us_totals <- us_counties %>% 
  group_by(state) %>% 
  summarise(total_cases = sum(cases),
            total_deaths = sum(deaths)) %>% 
  arrange(desc(total_deaths))
```

```{r Q1 View data frame}
# First 6 rows of `us_totals`
us_totals %>% 
  head()
```

# Question 2:
Get county names for DC, Maryland, and VA.
```{r Q2 County names, message=FALSE}
# Vector of country names
dmv_counties <- c("Anne Arundel", "Charles", "Frederick", "Howard", "Montgomery", "Prince George's", "District of Columbia", "Alexandria city", "Arlington", "Fairfax", "Loudoun", "Prince William", "Falls Church city")

# Vector of state names
dmv_states <- c("District of Columbia", "Maryland", "Virginia")
```

```{r Q2 Subset to DMV, message=FALSE}
# Subsetting to states and counties only in DMV
dmv <- us_counties %>% 
  # Filters to states in counties from vectors above
  filter(state %in% dmv_states) %>% 
  filter(county %in% dmv_counties) %>% 
  # Getting rid of duplicate counties:
    ## Montgomery VA fips = 51121
    ## Frederick VA fips = 51069
  filter(fips != "51121") %>% 
  filter(fips != "51069")
```

```{r Q2 DMV df, message=FALSE}
# DMV counties only, total cases and deaths, arranged by deaths
dmv_totals <- dmv %>% 
  group_by(county) %>% 
  summarise(total_cases = sum(cases),
            total_deaths = sum(deaths)) %>% 
  arrange(desc(total_deaths))
```

```{r Q2 DMV}
# First 6 rows of `dmv_totals`
dmv_totals %>% 
  head()
```

# Question 3:
Make a plot of the total number of cases by date, and the total of deaths by date, for counties in the DMV area. *Make sure to use facets and aesthetics* to convey results, and that the plot is titled, axes are labeled, and text is legible.
```{r Q3 DMV long}
# Lengthening data to make facets in graph
dmv_line <- dmv %>% 
  pivot_longer(cols = c(cases, deaths), names_to = "type", values_to = "count")
```

```{r Q3 DMV inital plot}
# Initial plot
dmv_line_plot <- dmv_line %>% 
  ggplot(aes(x = date, y = count, color = county)) +
  geom_line() +
  facet_wrap( ~ type, scales = "free_y") +
  theme_minimal()
```

```{r Q3 DMV final plot}
# Adding title, axes, other labels
dmv_line_plot + 
  labs(title = "Total Number of Cases and Deaths Per Day in the DMV Area",
       subtitle = "March through December 2020",
       x = "Date",
       y = "Count")
```

# Question 4:
Read in and tidy the global and US COVID-19 data contained in the four fields below:

  + "time_series_covid19_confirmed_global.csv"
  + "time_series_covid19_deaths_global.csv"
  + "time_series_covid19_confirmed_US.csv"
  + "time_series_covid19_deaths_US.csv"
  
```{r Q4 Loading data, message=FALSE}
confirmed_global <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")
deaths_global <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")
confirmed_US <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv")
deaths_US <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv")
```
  
For the US data, *produce a second* tidy data frame called `US_by_state` that has the total of confirmed cases, deaths, population, and deaths per million people for each date for each state.

```{r Q4 Removing old objects}
# Making space for memory
rm(dmv, dmv_line, dmv_line_plot, dmv_totals, us_totals)
```


```{r Q4 Joining US dfs, message=FALSE}
# Lengthening US cases
long_cases_US <- confirmed_US %>% 
  pivot_longer(cols = ends_with("20"), names_to = "date", values_to = "cases_count") %>% 
  group_by(Province_State, date) %>% 
  summarise(daily_case_count = sum(cases_count))  %>% 
  arrange(date)

# Lengthening US deaths
long_deaths_US <- deaths_US %>% 
  pivot_longer(cols = ends_with("20"), names_to = "date", values_to = "deaths_count") %>% 
  group_by(Province_State, date) %>% 
  summarise(daily_deaths_count = sum(deaths_count), population = sum(Population)) %>% 
  arrange(date)

# Joining tables
US_by_state <- long_cases_US %>% 
  left_join(long_deaths_US, by = "Province_State") %>% 
  arrange(date.x, date.y)

# Fixing order of columns and renaming columns
US_by_state <- US_by_state %>% 
  rename(location = Province_State,
         date = date.x) %>% 
  select(location, population, date, daily_case_count, daily_deaths_count,-date.y)

# Adding deaths per million column and dropping NA's
US_by_state <- US_by_state %>% 
  mutate(deaths_millions = (daily_deaths_count / population) * 1000000) %>% 
  drop_na() %>% 
  filter(location != "Grand Princess")
```

```{r}
US_by_state %>% 
  head()
```


```{r}
# Cleaning up memory
rm(long_cases_US, long_deaths_US, confirmed_US, deaths_US)
```


# Questoin 5:
For the global data, create a new tidy data frame called `global` that has the total confirmed cases and deaths per country by data. Add global population data and replace the US observations in `global` with the US data.

```{r Q5 Loading country info}
# Reading new global data
country_info <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv")
```

```{r Q5 US pivot longer, message=FALSE}
# Lengthening US cases
long_cases_global <- confirmed_global %>% 
  pivot_longer(cols = ends_with("20"), names_to = "date", values_to = "cases_count") %>% 
  group_by(`Country/Region`, date) %>% 
  summarise(daily_case_count = sum(cases_count))  %>% 
  arrange(date) %>% 
  rename(country_region = `Country/Region`)

# Joining population data
deaths_global <- deaths_global %>% 
  left_join(country_info, by = "Lat")

# Lengthening US deaths
long_deaths_global <- deaths_global %>% 
  pivot_longer(cols = ends_with("20"), names_to = "date", values_to = "deaths_count") %>% 
  group_by(`Country/Region`, date) %>% 
  summarise(daily_deaths_count = sum(deaths_count), population = sum(Population)) %>% 
  arrange(date) %>% 
  rename(country_region = `Country/Region`)
```

```{r Q5 Global joins}
# Creating global dataframe
global <- long_deaths_global %>% 
  left_join(long_cases_global, by = "country_region") %>% 
  arrange(date.x, date.y)

# Fixing order of columns and renaming columns
global <- global %>% 
  rename(date = date.x,
         location = country_region) %>% 
  select(location, population, date, daily_case_count, daily_deaths_count,-date.y)

# Adding deaths per million column
global <- global %>% 
  mutate(deaths_millions = (daily_deaths_count / population) * 1000000) %>% 
  drop_na()
```

```{r Q5 Adding state data}
# Replace US data
global <- global %>% 
  filter(location != "US")

# Adding US states
global <- global %>% 
  bind_rows(US_by_state)

# Correcting year
global$date <- paste(global$date, "20", sep = "")

# Converting to date
global$date <- as_date(global$date, format = "%m/%d/%Y")
```


```{r}
# Making space in memory
rm(confirmed_global, deaths_global, long_cases_global, long_deaths_global, deaths_global, country_info)
```


# Question 6:
Create a new variable in `global` representing the deaths per million per country by date, and remove any rows with no confirmed cases.

Included in code above.

```{r Global df}
global %>% 
  head()
```


# Question 7:
Find the 25 countries or US states with the most confirmed cases (**use `group_by()`, `summarise()`, and `ungroup()`**). What are some possible explanations for the difference in number of deaths per million people among these groups?

```{r Q7 Top 25 countries}
# Creating data frame with top 25 countries
top_countries <-  global %>% 
  group_by(location) %>% 
  summarise(total_case_count = sum(daily_case_count),
            total_death_millions = sum(deaths_millions)) %>% 
  ungroup() %>%
  arrange(desc(total_case_count)) %>% 
  head(25)

top_countries
```
Some potential explanations for high confirmed cases of COVID-19 could be due to the country or state having a large population and low restrictions on social distancing. The differences between case count and number of deaths per millions could be the result of varying qualities of healthcare infrastructure. 

# Question 8:
For the 25 countries/US states with the highest number of confirmed cases, construct plots for:

  1. Confirmed cases over time
  2. Deaths over time
  3. Deaths per million over time
  
**Facet each plot by continent** and use appropriate scales for the axes.
```{r Q8 Adding continent}
# Saving vector of country names for reference
countries <- top_countries$location

# Subset of global data for graph, adding continent using `case_when()`
global_line <- global %>% 
  filter(location %in% countries) %>% 
   mutate(continent = case_when(
    location %in% c("Texas", "California", "Florida", "New York", "Illinois", "Georgia", "Mexico") ~ "North America",
    location %in% c("Brazil", "Peru", "Colombia", "Argentina", "Chile") ~ "South America",
    location %in% c("Spain", "United Kingdom", "Germany", "Italy") ~ "Europe",
    location == "South Africa" ~ "Africa",
    location %in% c("India", "Iran", "Turkey", "Bangladesh", "Saudi Arabia", "Pakistan", "Iraq", "Russia") ~ "Asia")) %>%
  drop_na()
```

```{r}
# Memory
rm(top_countries)
```


```{r}
# Convert to date
global_line$date <- as_date(global_line$date, format = "%m/%d/%Y")

# Grouping by continent
global_line <- global_line %>% 
  group_by(continent, date) %>% 
  summarize(daily_case_count = sum(daily_case_count),
            daily_deaths_count = sum(daily_deaths_count),
            deaths_millions = sum(deaths_millions))
```


## Confirmed Cases over time
```{r Q8 Cases plot}
# Initial Plot
cases_line_plot <- global_line %>% 
  ggplot(aes(x = date, y = daily_case_count)) +
  geom_line() +
  facet_wrap( ~ continent, scales = "free") +
  theme_minimal()

cases_line_plot <- cases_line_plot +
  labs(title = "Confirmed COVID-19 Cases Over Time",
       subtitle = "2020",
       x = "Date",
       y = "Confirmed Cases")

cases_line_plot
```

## Deaths over time
```{r Q8 Death plot}
# Initial Plot
deaths_line_plot <- global_line %>% 
  ggplot(aes(x = date, y = daily_deaths_count, color = continent)) +
  geom_line() +
  facet_wrap( ~ continent, scales = "free") +
  theme_minimal() +
  theme(legend.position = "none")

deaths_line_plot <- deaths_line_plot +
  labs(title = "Confirmed Deaths from COVID-19 Cases Over Time",
       subtitle = "2020",
       x = "Date",
       y = "Deaths")

deaths_line_plot
```

## Deaths per million over time
```{r Q8 Deaths per million plot}
# Initial Plot
millions_line_plot <- global_line %>% 
  ggplot(aes(x = date, y = deaths_millions, color = continent)) +
  geom_line() +
  facet_wrap( ~ continent, scales = "free") +
  theme_minimal() +
  theme(legend.position = "none")

millions_line_plot <- millions_line_plot +
  labs(title = "Confirmed Deaths in Millions from COVID-19 Cases Over Time",
       subtitle = "2020",
       x = "Date",
       y = "Deaths in Millions")

millions_line_plot
```


# Question 9:
Get the data for the state of NY from the csse dataset and compare the growth of cases to the growth rate shown in the nytimes data. Does the data match? If not, how is it different?
```{r Q9 Comparing NY data}
# Growth for NYT data
nyt_ny <- us_counties %>% 
  filter(state == "New York") %>% 
  group_by(state, date) %>% 
  summarise(cases = cumsum(cases))

# Plot of New York Times data
nyt_ny %>% 
  ggplot(aes(x = date, y = cases)) +
  geom_line(color = "light pink") +
  theme_minimal() +
  labs(title = "New York Cases Counts",
       subtitle = "Source: New York Times",
       x = "Date",
       y = "Number of Cases")

# Growth for CSSE data
csse_ny <- US_by_state %>% 
  filter(location == "New York")

# Formatting date column
csse_ny$date <- paste(csse_ny$date, 20, sep = "")
csse_ny$date <- as_date(csse_ny$date, format = "%m/%d/%Y")

# Calculating growth
csse_ny <- csse_ny %>% 
  group_by(location, date) %>% 
  summarise(cases = cumsum(daily_case_count))

# Plot of CSSE data
csse_ny %>% 
  ggplot(aes(x = date, y = cases)) +
  geom_line(color = "hot pink") +
  theme_minimal() +
  labs(title = "New York Cases Counts",
       subtitle = "Source: CSSE",
       x = "Date",
       y = "Number of Cases")
```
The plots of the two data have an overall similar shape, but the CSSE has calculated a much higher number of cases.

# Question 10:
Plot the number of confirmed cases, deaths, and deaths per million inhabitants for Sweden, Denmark, and Finland.

## Sweden
```{r Q10 Sweden df}
sweden <- global %>% 
  filter(location == "Sweden")
```

```{r Sweden plot 1}
sweden %>% 
  ggplot(aes(x = daily_case_count)) +
  geom_boxplot(fill = "light blue") +
  theme_minimal() +
  labs(title = "Distribution of Daily COVID-19 Case Counts in Sweden",
       subtitle = "2020",
       x = "Case Count")
```

```{r Sweden plot 2}
sweden %>% 
  ggplot(aes(x = date, y = daily_deaths_count)) +
  geom_line(color = "light blue") +
  theme_minimal() +
  labs(title = "Counts of Daily Deaths due to COVID-19 in Sweden",
       subtitle = "2020",
       x = "Date",
       y = "Count")
```

```{r Sweden plot 3}
sweden %>% 
  ggplot(aes(x = date, y = deaths_millions)) +
  geom_line(color = "light blue") +
  theme_minimal() +
  labs(title = "Daily Deaths in Millions due to COVID-19 in Sweden",
       subtitle = "2020",
       x = "Date",
       y = "Count in Millions")
```


## Denmark
```{r Denmark df}
denmark <- global %>% 
  filter(location == "Denmark")
```

```{r Denmark plot 1}
denmark %>% 
  ggplot(aes(x = daily_case_count)) +
  geom_boxplot(fill = "light yellow") +
  theme_minimal() +
  labs(title = "Distribution of Daily COVID-19 Case Counts in Denmark",
       subtitle = "2020",
       x = "Case Count")
```

```{r Denmark plot 2}
denmark %>% 
  ggplot(aes(x = date, y = daily_deaths_count)) +
  geom_line(color = "orange") +
  theme_minimal() +
  labs(title = "Counts of Daily Deaths due to COVID-19 in Denmark",
       subtitle = "2020",
       x = "Date",
       y = "Count")
```

```{r Denmark plot 3}
denmark %>% 
  ggplot(aes(x = date, y = deaths_millions)) +
  geom_line(color = "Orange") +
  theme_minimal() +
  labs(title = "Daily Deaths in Millions due to COVID-19 in Denmark",
       subtitle = "2020",
       x = "Date",
       y = "Count in Millions")
```


## Finland
```{r Finland df}
finland <- global %>% 
  filter(location == "Finland")
```

```{r Finland plot 1}
finland %>% 
  ggplot(aes(x = daily_case_count)) +
  geom_boxplot(fill = "light green") +
  theme_minimal() +
  labs(title = "Distribution of Daily COVID-19 Case Counts in Finland",
       subtitle = "2020",
       x = "Case Count")
```

```{r Finland plot 2}
finland %>% 
  ggplot(aes(x = date, y = daily_deaths_count)) +
  geom_line(color = "light green") +
  theme_minimal() +
  labs(title = "Counts of Daily Deaths due to COVID-19 in Finland",
       subtitle = "2020",
       x = "Date",
       y = "Count")
```

```{r Finland plot 3}
finland %>% 
  ggplot(aes(x = date, y = deaths_millions)) +
  geom_line(color = "light green") +
  theme_minimal() +
  labs(title = "Daily Deaths in Millions due to COVID-19 in Finland",
       subtitle = "2020",
       x = "Date",
       y = "Count in Millions")
```

```{r}
rm(finland, denmark, sweden, temp)
```


How do the number of deaths per million residents compare? What are some possible explanations for any differences you observe?

It appears that the deaths per million increased at the fastest rate in April and May for Finland and Denmark. Whereas, in Sweden, the growth rate was more gradual and lasted mostly from Aprily through July. However, the overall number of people dying is higher overall for Sweden. The differences between these countries could be explained by their respective governments response to handling social distancing and the size of the countries populations.
