---
title: "Exercise 1"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r, message=FALSE}
library(tidyverse)
```


# Exercise 1
```{r}
new_string <- "Jane says \"thanks!\""

print(new_string)

writeLines(new_string)

writeLines("\\\"\\")
```

# Exercise 2
## `str_length()` - get length
```{r}
exercise2_tbl <- tribble(
                         ~Name, ~Age, 
                         "Alice", 25,
                         "Bob",   30,
                         "Chris", 35
)

exercise2_tbl <- exercise2_tbl %>% 
  mutate(value = Age * str_length(Name)) %>% 
  arrange(desc(value))

exercise2_tbl
```

# Exercise 3
## `str_c` - concatenate
```{r}
area_codes <- c(703, 863, 404, 202)
phone_nums <- c(5551212, 1234567, 7891234, 4799747)

str_c("(", area_codes, ") ", phone_nums, sep = "")
```

# Exercise 4
## `str_sub()` - extract character subsets
## `str_trim()` - removes whitespace from start and/or end
## `str_squish()` - reduces repeated whitespace inside a string

```{r}
x <- "    Help me to     get rid     of extra spaces?      "

str_trim(x, side = "left")
str_trim(x, side = "right")
str_trim(x, side = "both")

str_squish(x)
```

# Exercise 5
## Use `str_length()` and `str_sub()` to extract the middle character from a string.

```{r}
middle_character <- function(word) {
  
  # Get length of word
  len <- str_length(word)
  
  # Divide by two to find middle letter
  middle <- (len + 1) / 2
  even_mid <- (len) / 2 + 1
  
  # Case for even number
  if (len %% 2 == 0) {
    str_sub(word, middle, even_mid)
  }
  else {
    # Extract middle character
  str_sub(word, middle, middle)
  }
}
```

```{r}
middle_character("hamburger")
middle_character("hotdog")
```

# 3. Matching Patterns
## `str_view()` and `str_view_all()` - find a given pattern
# Exercise 6
Match this sequence: "'\?

```{r}
s <- str_c("some stuff", "\"'\\?", "more!")
```

```{r}
str_view(s, "\"\'\\\\.")
```

# Exercise 7
## ^ - starts with
## $ - end with
## ^start and end with$
Create regular expressions that find all words that start with "y" and end with "x" using `stringr::words`.

```{r}
words %>% 
  str_view_all("^y", match = TRUE)

words %>% 
  str_view_all("x$", match = TRUE)
```

