---
title: 'New York City SAT Scores: Exploratory Data Analysis'
author: "Carmen Canedo"
date: "28 September 2020"
output: pdf_document
---

```{r, include=FALSE}
# Global option to hide all code but still show results
knitr::opts_chunk$set(echo = FALSE)
```


```{r, message=FALSE, warning=FALSE}
# Accessing R file with our functions
source("setup.R")

# Creating new subdirectories
create_subdirectories(c("Data", "Results"))

# Reading in and cleaning the data
SAT_tbl <- get_data("NYC_SAT_Data.csv",
                    "Data",
                    "https://data.cityofnewyork.us/api/views/zt9s-n5aj/rows.csv?accessType=DOWNLOAD") %>% 
  transform_data()
```

## 1. & 2. Box plots of SAT distributions by borough
```{r}
# Plotting reading boxplot
box_reading <- SAT_tbl %>% 
  # Setting y axis as critical reading mean
  ggplot(aes(x = 1,
             y = Critical_Reading_Mean)) +
  # Creating bocplot and setting colors
  geom_boxplot(fill = "#ffe380",
               color = "#e6c345",
               alpha = 0.8) +
  # Faceting by Borough
  facet_grid(~ Borough) +
  # Adjusting y axis scal
  scale_y_continuous(breaks = seq(200, 800, 50)) +
  # Setting theme
  theme_bw() +
  # Getting rid of text on x axis, changing color of facet rectangles, and changing font
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_rect(fill = "#7dc760"),
        text = element_text(family = "serif")) +
  # Writing out labels
  labs(y = "Average Crtitical Reading Scores",
       title = "Distribution of Average Critical Reading Scores for the SAT by NYC Borough",
       caption = "Source: https://data.cityofnewyork.us/")

# Plotting reading boxplot
box_math <- SAT_tbl %>% 
  ggplot(aes(y = Mathematics_Mean)) +
  geom_boxplot(fill = "#809cff",
               color = "#4568e6",
               alpha = 0.8) +
  facet_grid(~ Borough) +
  scale_y_continuous(breaks = seq(200, 800, 50)) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_rect(fill = "#7dc760"),
        text = element_text(family = "serif")) +
  labs(y = "Average Math Scores",
       title = "Distribution of Average Math Scores for the SAT by NYC Borough",
       caption = "Source: https://data.cityofnewyork.us/")

# Plotting reading boxplot
box_writing <- SAT_tbl %>% 
  ggplot(aes(y = Writing_Mean)) +
  geom_boxplot(fill = "#e380ff",
               color = "#c345e6",
               alpha = 0.8) +
  facet_grid(~ Borough) +
  scale_y_continuous(breaks = seq(200, 800, 50)) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_rect(fill = "#7dc760"),
        text = element_text(family = "serif")) +
  labs(y = "Average Writing Scores",
       title = "Distribution of Average Writing Scores for the SAT by NYC Borough",
       caption = "Source: https://data.cityofnewyork.us/")
```

```{r}
# Collect highest scoring schools
high_scores <- SAT_tbl %>% 
  arrange(desc(Critical_Reading_Mean, Mathematics_Mean, Writing_Mean)) %>% 
  distinct(Borough, .keep_all = TRUE)
```

```{r}
# Adding annotation of the schools with the highest mean scores
# Collection of schools and Boroughs to use as labels
data_text <- data.frame(
  label = c("Stuyvesant HS",
            "HS of American Studies",
            "Brooklyn Tech HS",
            "Townsend Harris HS",
            "Staten Island Tech"),
  Borough = c("Manhattan",
              "Bronx",
              "Brooklyn",
              "Queens",
              "Staten Island"))

# Adding label to boxplots created above
box_reading <- box_reading +
  geom_text(data = data_text,
            mapping = aes(x = Inf, y = Inf, label = label),
            hjust = 1.1,
            vjust = 1.5,
            size = 2.5) +
  theme(axis.title.x = element_blank())

box_math <- box_math +
    geom_text(data = data_text,
            mapping = aes(x = Inf, y = Inf, label = label),
            hjust = 1.1,
            vjust = 1.5,
            size = 2.5) +
  theme(axis.title.x = element_blank())

box_writing <- box_writing +
  geom_text(data = data_text,
            mapping = aes(x = Inf, y = Inf, label = label),
            hjust = 1.1,
            vjust = 1.5,
            size = 2.5) +
  theme(axis.title.x = element_blank())
```

```{r, fig.width=7, fig.height=3}
# Calling plots
box_reading
box_math
box_writing
```

## 3. Summary stats of scores for each section of the SAT in each borough
```{r message=FALSE}
# Calculating stats for reading scores
stats_reading <- SAT_tbl %>% 
  group_by(Borough) %>% 
  summarise(mean_reading = mean(Critical_Reading_Mean), median_reading = median(Critical_Reading_Mean), sd_reading = sd(Critical_Reading_Mean))

# Stats for math scores
stats_math <- SAT_tbl %>% 
  group_by(Borough) %>% 
  summarise(mean_math = mean(Mathematics_Mean), median_math = median(Mathematics_Mean), sd_math = sd(Mathematics_Mean))

# Stats for writing scores
stats_writing <- SAT_tbl %>% 
  group_by(Borough) %>% 
  summarise(mean_writing = mean(Writing_Mean), median_writing = median(Writing_Mean), sd_writing = sd(Writing_Mean))

# Joining all tables
interim <- left_join(stats_reading, stats_math, by = "Borough")
stats_total <- left_join(interim, stats_writing, by = "Borough")

# Rename columns =
stats_total <- stats_total %>% 
  rename(`Mean ` = mean_reading,
         `Median ` = median_reading,
         `SD ` = sd_reading,
         `Mean  ` = mean_math,
         `Median  ` = median_math,
         `SD  ` = sd_math,
         ` Mean ` = mean_writing,
         ` Median ` = median_writing,
         ` SD ` = sd_writing)
```

```{r}
# Creating table with summary statistics
kable(stats_total) %>% 
  add_header_above(c(" " = 1, "Reading Scores" = 3, "Math Scores" = 3, "Writing Scores" = 3))
```


## 4. Exploring the relationship between the number of test takers and SAT scores for each section and across the five boroughs.
```{r}
# Creating scatterplot for reading scores
viz_reading_avg <- SAT_tbl %>% 
  ggplot(aes(x = Number_of_Test_Takers, y = Critical_Reading_Mean, color = Borough)) +
  geom_jitter() +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 1500, 150)) +
  scale_y_continuous(breaks = seq(200, 800, 50)) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  labs(x = "Number of Test Takers",
       y = "Average Reading Scores",
       title = "Average Reading Scores for the SAT by Number of Test Takers in NYC",
       caption = "Source: https://data.cityofnewyork.us/")

# Calling plot
viz_reading_avg

# Scatterplot for math
viz_math_avg <- SAT_tbl %>% 
  ggplot(aes(x = Number_of_Test_Takers, y = Mathematics_Mean, color = Borough)) +
  geom_jitter() +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 1500, 150)) +
  scale_y_continuous(breaks = seq(200, 800, 50)) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  labs(x = "Number of Test Takers",
       y = "Average Math Scores",
       title = "Average Math Scores for the SAT by Number of Test Takers in NYC",
       caption = "Source: https://data.cityofnewyork.us/")

# Calling plot
viz_math_avg

# Scatterplot for writing
viz_writing_avg <- SAT_tbl %>% 
  ggplot(aes(x = Number_of_Test_Takers, y = Writing_Mean, color = Borough)) +
  geom_jitter() +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 1500, 150)) +
  scale_y_continuous(breaks = seq(200, 800, 50)) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  labs(x = "Number of Test Takers",
       y = "Average Writing Scores",
       title = "Average Writing Scores for the SAT by Number of Test Takers in NYC",
       caption = "Source: https://data.cityofnewyork.us/")

# Calling plot
viz_writing_avg
```


## 5. Results of analysis
In my analysis, I have found that Staten Island schools tended to preform higher on individual sections of the ACT, but Manhattan had the greatest variance and the highest scores came from Stuyvesant. It appears that there is not much correlation between the number of test takers and the average scores by subject. Perhaps in a further study I can examine factors such as age, race, or socio-economic backgrounds to see if this has any influence on test results.