---
title: "Week 3 Survey"
author: "Carmen Canedo"
date: "8 September 2020"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: inline
---

# Loading libraries
```{r, message=FALSE}
library(tidyverse)
```

## 1. Assign the value 10 to variable x
```{r}
x <- 10
```

## 2. Assign names to variable called names
This is a character vector.
```{r}
names <- c("Alex", "Bob", "Chris", "Donna")
```

## 3. Whic value is greater?
The first expression is greater.
```{r}
(log10(5^5) / 2)

pi / 2
```

## 4. Storing logical value
```{r}
test <- (log10(5^5) / 2) > pi / 2
```

## 5. Descriibe differences in data stypes
  * **Scalars** are just a single variable
  * **Vectors** are a collection of items of the same type - similar to lists but they can only hold one class
  * **Lists** are a collection of any type of objects
  * **Matrices** are two dimensional collections of data all of the same class
  * **Data frames** are also two dimensional but can hold any type of data 
  
## 6. Create a list
```{r}
collection <- list(x, names, test)
collection
```

## 7. Create data frame
```{r}
# Making vectors of the data
new_names <- c("Alex", "Bob", "Chris", "Donna")
age <- c(25, 21, 29, 32)
likes_pizza <- c(TRUE, TRUE, FALSE, FALSE)

# Creating tibble
pizza <- tibble(name = new_names, Age = age, likes_pizza = likes_pizza)
pizza
```

## 8. Fix column name and add column
```{r}
# Collecting height data
height = c(65, 72, 74, 68)

pizza <- pizza %>% 
  rename(age = Age) %>% 
  mutate(height = height)

pizza
```

## 9. Create function called marketability
```{r}
marketability <- function(age, height, likes_pizza = TRUE) {
  if (likes_pizza == TRUE) {
    score <- (height / 60) * (height / 55)^2
    score
  } else {
    score <- (height / 15) * (height / 55)^2
    score
  }
}
```

## 10. Which individuals shouldn't be targets of the campagin?
```{r}
# Calculating score
alex <- marketability(25, 65)
bob <- marketability(21, 72)
chris <- marketability(29, 74, likes_pizza = FALSE)
donna <- marketability(32, 68, likes_pizza = FALSE)

# Checking to see if score is less than 6
alex < 6
bob < 6
chris < 6
donna < 6
```
Because Alex and Bob's marketability scores are less than 6, they should not be targets of the campaign.
