---
title: "Homework 5"
author: "Carmen Canedo"
date: "2 October 2020"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: inline
---

```{r, include=FALSE}
# Loading libraries via R file
source("setup.R")
```


## Exercise 2.1 Why does tidy data lend itself to vectorised operations?
Tidy data ensures that an observation is always correctly paired with the variables

## Exercise 2.2 How could you tidy the SAT data from last week? Which of the data sets below are tidy? What's wrong with the non-tidy data sets?
After reading in the SAT data from the .csv, I placed my data into a tibble. Each variable is in a column, each observation has its own row, and each value has its own cell. To make the SAT data tidy, I would also make all variable names names use snake case (all lowercase letters and underscore instead of spaces), and use title case for all the name of the high schools for consistency.

The only table that is tidy:

 + Table 1
 
The following **are not** tidy:

  + Table 2
    + `rate` contains two variables. To fix this we can separate them into `num_cases` and `total_population`, and if we still wanted to include the rate, we could use `mutate()` to divide the two and store the values in `rate`.
    
  + Table 3
    + `2000` and `1999` belong to one variable `year`, but in this table, they are spread across two columns. To fix this, we can use `pivot_longer()` and assign column names to `year` and the values to a separate column, `num_cases`.
    
  + Table 4
    + The observations (country names) in the rows are repeated, so we can use `pivot_wider()` to split `type` into `num_cases` and `total_population`
    

## Exercise 2.3 Use `pivot_longer()` to tidy data frame
```{r}
# Looking at data frame
relig_income %>%
  head(n = 6)
```

```{r}
# Pivoting longer, using all columns but religion
relig_income %>% 
  pivot_longer(cols = !religion, names_to = "income", values_to = "count")
```


## Exercise 2.4 Tidy the data from blackboard
```{r}
# Reading in data
monkey <- read_csv("monkey_mem.csv")

# Pivoting longer
monkey %>% 
  pivot_longer(cols = starts_with("W"), names_to = "week", values_to = "accuracy")
```


## Exercise 2.5 Use `pivot_wider()` to tidy `tidyr::fishencounters`
```{r}
# Pivoting wider
fish_encounters %>% 
  pivot_wider(names_from = station, values_from = seen)
```

## Exercise 2.6 Tidy flowers1 data set
```{r}
#Reading in flowers data
flowers1 <- read_delim("flowers1.csv", delim = ";")
```

```{r}
# Pivoting wider
flowers1 %>% 
  pivot_wider(names_from = Variable, values_from = Value)
```


## Exercise 2.7 Use `separate` to tidy the flowers2 data set
```{r}
# Reading in flowers 2 data
flowers2 <- read_delim("flowers2.csv", delim = ";")
```

```{r}
# Getting rid of /
flowers2 %>% 
  separate(`Flowers/Intensity`,
           into = c("flowers", "intensity"),
           sep = "/",
           convert = TRUE)
```


## Exercise 2.8 Read the help file for `unite` and correct the code above to get rid of underscore in `year` column
```{r}
# Getting rid of underscore
table5 %>% 
  unite("year",
        c(century, year),
        sep = "")
```

## Exercise 2.9 Turn implicit missing values explicit in the data frame
```{r}
# Creating tibble
output <- tibble(treatment = c("a", "b", "a", "c", "b"),
                 gender = factor(c("M", "F", "F", "M", "M"),
                                 levels = c("M", "F", "O")),
                 return = c(1.5, 0.75, 0.5, 1.8, NA)
)

output
```

```{r}
# Making missing values explicit
output %>% 
  complete(treatment, nesting(gender, return ))
```


## Exercise 2.10 Tidy the `tidyr::billboard` data set
### 1: Gather up all the week entries into a row for each week for each song where there is an entry
```{r}
# Pivoting longer and renaming date entered
billboard_tidy <- billboard %>% 
  rename("date_entered" = "date.entered") %>% 
  pivot_longer(starts_with("w"), names_to = "week", values_to = "position")
```

### 2: Convert the week variable to a number and figure out the date corresponding to each week on the chart
```{r}
# Parsing numbers from week values, converting date, getting rid of date_entered
billboard_tidy <- billboard_tidy %>% 
  mutate(week = parse_number(week),
         date = as.Date(date_entered) + 7 * (week - 1)) %>% 
  select(-date_entered)
```

### 3: Sort the data by artist, track, and week
```{r}
# Sorting data
billboard_tidy <- billboard_tidy %>% 
  drop_na() %>% 
  arrange(artist, track, week)
```

```{r}
# Output
billboard_tidy
```

